Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO$_2$ Reduction Reactions

### 2024-01-04 paper/main_sections/1_head.tex
temp save: format cleanup

Authors
Haoyu Yang, Juanli Zhao, Qiankun Wang, Bin Lin, Wei Luo, Ziqi Sun*, Ting Liao*

Title
Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO2 Reduction Reactions

>>>>

% Title and Authors (Affiliations)

% Title
\title{Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO$_2$ Reduction Reactions}

% Authors and Affiliations
\author[1]{Haoyu Yang}
\author[3]{Juanli Zhao}
\author[4]{Qiankun Wang}
\author[5]{Bin Lin}
\author[6]{Wei Luo}
\author[1,*]{Ziqi Sun}
\author[2,**]{Ting Liao}

\affil[1]{School of Chemistry and Physics, Queensland University of Technology, George Street, Brisbane, QLD 4000, Australia}
\affil[2]{School of Mechanical Medical and Process Engineering, Queensland University of Technology, George Street, Brisbane, QLD 4000, Australia}
\affil[3]{School of Materials Science and Engineering, Shanghai University, Shanghai, China}
\affil[4]{School of Computer Science and Engineering, Southeast University, Nanjing, China}
\affil[5]{State Key Laboratory for Modification of Chemical Fibers and Polymer Materials, College of Materials Science and Engineering, Institute of Functional Materials, Donghua University, Shanghai, 201620 China}
\affil[*]{Corresponding author: ziqi.sun@qut.edu.au}
\affil[**]{Corresponding author: t3.liao@qut.edu.au}

% Date (optional)
\date{2024-01-04}

% Make Title
\maketitle

### 2024-01-04 paper/main_sections/2_abstract.tex
temp save: format cleanup

Abstract
Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions, yet the diverse combinations of active elements and support materials, the nature of coordination environments, elude traditional methodologies in searching optimal SAC systems with superior catalytic performance. Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models to hybrid descriptor based activity volcano plot, two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports, including g-C3N4, nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, BN, and C2N, are screened for efficient CO2 reduction reaction (CO2RR). Starting from establishing a correlation map between the adsorption energies of intermediates and diverse electronic and elementary descriptors, sole singular descriptor lost magic to predict catalytic activity, including d-band center. Deep learning method utilizing multi-branch CNN model therefore was employed, using 2D electronic density of states (eDOS) as input to predict adsorption energies. Hybrid-descriptor enveloping both C- and O-types of CO2RR intermediates was introduced to construct volcano plots and limiting potential periodic table, aiming for intuitive screening of catalyst candidates for efficient CO2 reduction to CH4. The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy. To explore the electronic scale principle governing practical engineering catalytic CO2RR activity, orbitalwise eDOS shifting experiments based on CNN model were employed. The study involves examining the adsorption energy and, consequently, catalytic activities while varying supported single atoms. This work offers a tangible framework to inform both theoretical screening and experimental synthesis, thereby paving the way for systematically designing efficient SACs.

>>>>

% Abstract

% TODO: prevent words from being broken in the middle

\newpage

\begin{abstract}
Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions, yet the diverse combinations of active elements and support materials, the nature of coordination environments, elude traditional methodologies in searching optimal SAC systems with superior catalytic performance.
Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models to hybrid descriptor based activity volcano plot, two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports, including g-C$_3$N$_4$, nitrogen-doped graphene, graphene with dual-vacancy, black phosphorous, BN, and C$_2$N, are screened for efficient CO$_2$ reduction reaction (CO$_2$RR).
Starting from establishing a correlation map between the adsorption energies of intermediates and diverse electronic and elementary descriptors, sole singular descriptor lost magic to predict catalytic activity, including d-band center. Deep learning method utilizing multi-branch CNN model therefore was employed, using 2D electronic density of states (eDOS) as input to predict adsorption energies.
Hybrid-descriptor enveloping both C- and O-types of CO$_2$RR intermediates was introduced to construct volcano plots and limiting potential periodic table, aiming for intuitive screening of catalyst candidates for efficient CO$_2$ reduction to CH$_4$.
The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy.
To explore the electronic scale principle governing practical engineering catalytic CO$_2$RR activity, orbitalwise eDOS shifting experiments based on CNN model were employed.
The study involves examining the adsorption energy and, consequently, catalytic activities while varying supported single atoms.
This work offers a tangible framework to inform both theoretical screening and experimental synthesis, thereby paving the way for systematically designing efficient SACs.
\end{abstract}

### 2024-01-04 paper/main_sections/3_introduction.tex
temp save: format cleanup

Electrochemical CO2RR stands as a promising avenue for long-term seasonal energy storage [1] and mitigating excess CO2 in the atmosphere.

Despite decades of research, metallic copper remains a leading CO2RR electrocatalyst for hydrocarbon production [2], albeit burdened by substantial overpotential and low faradic efficiency and poor operational stability [3, 4]. In this regard, supported SACs have displayed remarkable activity and selectivity across various catalytic reactions [5, 6], leveraging their unique electronic properties and coordination environment. To date, experimental studies have primarily focused on two-electron reduction products due to the facile release of gaseous CO [7-9]. Consequently, there is a pressing need to modulate the electronic structure of single-atom sites to enable hydrocarbon production at reduced overpotentials.

@@

In this study, we introduced a catalyst performance analysis pipeline (Figure 1) integrating a customized CNN model and volcano plots. Our CNN model accurately estimated the adsorption energies, the activity descriptor, from the eDOS of 2D materials supported single metal catalysts. Evaluations on the adsorption energies of nine intermediates along CO2RR and competing hydrogen evolution reaction (HER) as well yielded prediction mean absolute errors (MAEs) at the level of 0.1 eV, which is well aligned with the precision of conventional DFT calculations [23-25]. This result validated the capability of current CNN model to accurately capture intricate spatial information from eDOS. Building on this, we introduced an orbital-wise eDOS occlusion experiment was adopted afterwards inspired from the occlusion technique widely used in computer vision and pattern recognition applications. Systematic eDOS experiment was performed to elucidate orbitalwise contributions to the adsorption energies at different energy levels, which produced consistent outcome from Crystal Orbital Hamilton Population (COHP) analysis and real-space wavefunction visualization. All these results justified the proficiency of CNN model on identifying the dominant orbitals for adsorbate-substrate interactions. Furthermore, orbital-wise eDOS shifting experiment, exploring the impact of orbital-wise eDOS shifting on the adsorption behavior of studied catalysts were performed.

Simultaneously, we incorporated the volcano plots into the catalysis performance analysis pipeline. Besides offering intuitive analysis of catalyst activities (represented by the elevation of volcanoes), volcano plots points to the directions on how effort can be exerted to identify promising catalyst candidates. Built upon scaling relations [26], we introduced a hybrid-descriptor scheme incorporating both C- and O-centered species. The scaling coefficient of determination (R2) was increased by 0.1387 using hybrid descriptor, which demonstrate an improved prediction accuracy compared to the well documented single descriptor based method.

In summary, we introduced a catalyst performance analysis pipeline by integrating CNN and volcano plot analysis, using CO2RR as an example proof of concept. The CNN, which correlates eDOS with activity descriptors, establishes a robust electronic structure-activity relationship. Notably, it is capable to predict the influence of eDOS disturbances, caused by either alloying, strain, or defects [27-29], to modulate the catalytic activity of materials, which will offer insights into the intrinsic mechanism for rationalized catalyst optimization strategies. Advancing the current volcano plot method, the involvement of CNN models in catalyst performance analysis pipeline holds enormous promise for advanced exploration of catalysts with superior performance.

>>>>

% Introduction

\newpage

\section{Introduction}
Electrochemical CO$_2$RR stands as a promising avenue for long-term seasonal energy storage [1] and mitigating excess CO$_2$ in the atmosphere.

Despite decades of research, metallic copper remains a leading CO$_2$RR electrocatalyst for hydrocarbon production [2], albeit burdened by substantial overpotential and low faradic efficiency and poor operational stability [3, 4].
In this regard, supported SACs have displayed remarkable activity and selectivity across various catalytic reactions [5, 6], leveraging their unique electronic properties and coordination environment.
To date, experimental studies have primarily focused on two-electron reduction products due to the facile release of gaseous CO [7-9].
Consequently, there is a pressing need to modulate the electronic structure of single-atom sites to enable hydrocarbon production at reduced overpotential.

@@

In this study, we introduced a catalyst performance analysis pipeline (Figure 1) integrating a customized CNN model and volcano plots.
Our CNN model accurately estimated the adsorption energies, the activity descriptor, from the eDOS of 2D materials supported single metal catalysts.
Evaluations on the adsorption energies of nine intermediates along CO$_2$RR and competing hydrogen evolution reaction (HER) as well yielded prediction mean absolute errors (MAEs) at the level of 0.1 eV, which is well aligned with the precision of conventional DFT calculations [23-25].
This result validated the capability of current CNN model to accurately capture intricate spatial information from eDOS.
Building on this, we introduced an orbital-wise eDOS occlusion experiment was adopted afterwards inspired from the occlusion technique widely used in computer vision and pattern recognition applications.
Systematic eDOS experiment was performed to elucidate orbitalwise contributions to the adsorption energies at different energy levels, which produced consistent outcome from Crystal Orbital Hamilton Population (COHP) analysis and real-space wavefunction visualization.
All these results justified the proficiency of CNN model on identifying the dominant orbitals for adsorbate-substrate interactions.
Furthermore, orbital-wise eDOS shifting experiment, exploring the impact of orbital-wise eDOS shifting on the adsorption behavior of studied catalysts were performed.

Simultaneously, we incorporated the volcano plots into the catalysis performance analysis pipeline.
Besides offering intuitive analysis of catalyst activities (represented by the elevation of volcanoes), volcano plots points to the directions on how effort can be exerted to identify promising catalyst candidates.
Built upon scaling relations [26], we introduced a hybrid-descriptor scheme incorporating both C- and O-centered species.
The scaling coefficient of determination ($R^2$) was increased by 0.1387 using hybrid descriptor, which demonstrate an improved prediction accuracy compared to the well documented single descriptor based method.

In summary, we introduced a catalyst performance analysis pipeline by integrating CNN and volcano plot analysis, using CO$_2$RR as an example proof of concept.
The CNN, which correlates eDOS with activity descriptors, establishes a robust electronic structure-activity relationship.
Notably, it is capable to predict the influence of eDOS disturbances, caused by either alloying, strain, or defects [27-29], to modulate the catalytic activity of materials, which will offer insights into the intrinsic mechanism for rationalized catalyst optimization strategies.
Advancing the current volcano plot method, the involvement of CNN models in catalyst performance analysis pipeline holds enormous promise for advanced exploration of catalysts with superior performance.

### 2024-01-04 paper/main_sections/4.1_cnn_for_eads.tex
temp save: format cleanup

Adsorption energy prediction from electronic density of states with convolutional neural network.

Originating from the pioneering work of Hammer and Nørskov [30], the d-band center model has emerged as a successful paradigm for employing descriptor to understand and predict the adsorption behavior on transition metal surfaces. To investigate alternative electronic and elementary descriptors that may serve as more suitable substitutes for the d-band center model in the context of SACs, we performed Kendall rank correlation analysis to account for both linearity and non-linearity [31]. The correlation coefficient map, as presented in Figure 2a, highlighted the d-band center as the most informative descriptor with a Kendall's τ coefficient of -0.33, and followed by the vacuum level and electronegativity. However, d-band center alone does not suffice for accurate predictions of intermediate adsorption energy as illustrated in the scatter plot Supplementary Figure 18, and therefore, refinements are imperative. Unlike bulk transition metal surfaces, where the variations in the d-band centre account for the majority of the adsorption energy variations [27, 32], the unique structure of SACs introduces more variables rather than d-band center that are intrinsically correlated with the variations of adsorption energy. This finding also interpreted that in some reported prior studies, indicating that depending solely on the d-band center may be insufficient for accurate adsorption energy predictions in the context of SACs [33-37]. As a result, the development of a method tailored specifically for catalysts like SACs becomes essential.

In our study, we employ CNN, a renowned deep learning method proficient in capturing spatial hierarchy from matrix-like data, such as the eDOS. Notably, CNN's autonomy from prior knowledge and human intervention during feature extraction renders it more efficient than traditional machine-learning methods. Our approach harnesses the entire 2D eDOS with a multi-branch CNN model. Figure 2b and Figure 2c delineate the model and its branches. The workflow involves processing the entire 2D eDOS as input, segregating orbitals into branches, and directing the concatenated outputs through another convolutional block, which generates the adsorption energy prediction. Sharing learnable parameters across branches optimizes computational efficiency. By leveraging the CNN model, we predicted adsorption energies from the eDOS of supported single atoms, and validated through parity plots (Figure 2d-m). The model predicts adsorption energies across nine species within the CO2RR process and competing HER side reaction, attaining an averaged MAE of 0.06 eV. This MAE aligns with the established precision of DFT methods, which typically range between 0.1-0.2 eV concerning experimental measurement [25, 38, 39]. The full list of prediction errors is available in Supplementary Table 18. These results confirm the CNN model's ability to predict adsorption energies from eDOS input. Moreover, they suggest the potential for predicting adsorption energies for SACs using eDOS, although this would demand more sophisticated feature engineering techniques.

>>>>

% Results and Discussion Part One: Adsorption energy prediction from eDOS with CNN

\section{Adsorption energy prediction from electronic density of states with convolutional neural network}
Originating from the pioneering work of Hammer and Nørskov [30], the d-band center model has emerged as a successful paradigm for employing descriptor to understand and predict the adsorption behavior on transition metal surfaces.
To investigate alternative electronic and elementary descriptors that may serve as more suitable substitutes for the d-band center model in the context of SACs, we performed Kendall rank correlation analysis to account for both linearity and non-linearity [31].
The correlation coefficient map, as presented in Figure 2a, highlighted the d-band center as the most informative descriptor with a Kendall's \tau coefficient of -0.33, and followed by the vacuum level and electronegativity.
However, d-band center alone does not suffice for accurate predictions of intermediate adsorption energy as illustrated in the scatter plot Supplementary Figure 18, and therefore, refinements are imperative.
Unlike bulk transition metal surfaces, where the variations in the d-band centre account for the majority of the adsorption energy variations [27, 32], the unique structure of SACs introduces more variables rather than d-band center that are intrinsically correlated with the variations of adsorption energy.
This finding also interpreted that in some reported prior studies, indicating that depending solely on the d-band center may be insufficient for accurate adsorption energy predictions in the context of SACs [33-37].
As a result, the development of a method tailored specifically for catalysts like SACs becomes essential.

In our study, we employ CNN, a renowned deep learning method proficient in capturing spatial hierarchy from matrix-like data, such as the eDOS.
Notably, CNN's autonomy from prior knowledge and human intervention during feature extraction renders it more efficient than traditional machine-learning methods.
Our approach harnesses the entire 2D eDOS with a multi-branch CNN model.
Figure 2b and Figure 2c delineate the model and its branches.
The workflow involves processing the entire 2D eDOS as input, segregating orbitals into branches, and directing the concatenated outputs through another convolutional block, which generates the adsorption energy prediction.
Sharing learnable parameters across branches optimizes computational efficiency.
By leveraging the CNN model, we predicted adsorption energies from the eDOS of supported single atoms, and validated through parity plots (Figure 2d-m).
The model predicts adsorption energies across nine species within the CO$_2$RR process and competing HER side reaction, attaining an averaged MAE of 0.06 eV.
This MAE aligns with the established precision of DFT methods, which typically range between 0.1-0.2 eV concerning experimental measurement [25, 38, 39].
The full list of prediction errors is available in Supplementary Table 18.
These results confirm the CNN model's ability to predict adsorption energies from eDOS input.
Moreover, they suggest the potential for predicting adsorption energies for SACs using eDOS, although this would demand more sophisticated feature engineering techniques.

### 2024-01-04 paper/main_sections/4.2_volcano_plots.tex
temp save: format cleanup

Hybrid descriptors enabled volcano plot for catalyst analysis.

Catalyst activity volcano plots, built upon the scaling relations proposed by Peterson and Nørskov's [26], goes beyond assessment of candidatures' energetics only but to provide informative guidance on discovering catalysts candidates with superior performance [42, 43]. Our endeavor to improve the accuracy of volcano plots leads to the development of the hybrid-descriptor, which encompasses both C- and O- bonding types of CO2RR intermediates within the scaling relations. This adaptation arose from the intuition that most CO2RR intermediates are featured with both bonding types rather than any single one in dominance. By integrating both bonding featured descriptors into the hybrid approach, we found the coefficient of determination (R2) was increased by 0.1387. Detailed information about the scaling parameters and performance enhancements are provided in Supplementary Tables 13 and Supplementary Figure 10. This improvement to some extent underscored our premise that CO2RR intermediates typically exhibit chemical similarities encompassing both O- and C-type descriptors. After taking this dual nature into consideration, our hybrid-descriptor method can evaluate the intermediates adsorption energies more accurately. Importantly, the introduction of hybrid descriptors is computationally affordable compared to DFT methods, as it involves only a set of linear regressions to identify the optimal mixing ratio, as elaborated in the Methods section.

Expanding upon the revised scaling relations, we introduce an activity volcano plot and a rate-determining step (RDS) volcano plot for the CO2RR process, shown in Figure 3a and Supplementary Figure 11. Given that HER is a key competing reaction with CO2RR [44], we also incorporated a selectivity volcano plot, presented in Supplementary Figure 12, to offer a comprehensive analysis of catalyst candidates. In adherence to Peterson and Nørskov's original framework [26], our choice of x and y axes descriptors aligned with Gads *CO and Gads *OH. Further elaboration on how the limiting potentials were evaluated across the volcano plot can be found in the Methods section. Among the candidatures explored, Ge@g-C3N4 emerged as the most promising catalyst, demonstrating theoretical activity at a limiting potential of -1.2126 eV. Notably, the observed selectivity trend paralleled the activity trend, suggesting that catalysts exhibiting high theoretical activity are also resistant to HER side reactions. The RDS volcano plot (Supplementary Figure 11) indicated protonation of *CO as the rate-determining step for nearly optimal candidates. Moreover, the activity volcano plot predicted an optimal catalyst featuring a theoretical limiting potential of -1.0516 V, at Gads *CO of 0.0589 eV and Gads *OH of -4.0075 eV, positioned at the right-center of the volcano plot. Results from RDS volcano plot indicate that protonation of *CO likely serves as the RDS for CO2RR. Consequently, it should be the focal point for future catalyst designs. The activity and selectivity volcano plots imply that enhancing theoretical activity could be achieved by decreasing CO affinity while increasing OH affinity with the catalysts. This result suggests a perspective: weak binding of CO to the catalyst surface might facilitate the rotation of C-O backbone from the “upright” configuration in *CO to the “horizontal” orientation in *CHO [45], and therefore facilitating dynamics and protonation of the *CO species.

>>>>

% Results and Discussion Part Two: Hybrid descriptors enabled volcano plot for catalyst analysis

\section{Hybrid descriptors enabled volcano plot for catalyst analysis}
Catalyst activity volcano plots, built upon the scaling relations proposed by Peterson and Nørskov's [26], goes beyond assessment of candidatures' energetics only but to provide informative guidance on discovering catalysts candidates with superior performance [42, 43].
Our endeavor to improve the accuracy of volcano plots leads to the development of the hybrid-descriptor, which encompasses both C- and O- bonding types of CO$_2$RR intermediates within the scaling relations.
This adaptation arose from the intuition that most CO$_2$RR intermediates are featured with both bonding types rather than any single one in dominance.
By integrating both bonding featured descriptors into the hybrid approach, we found the coefficient of determination ($R^2$) was increased by 0.1387.
Detailed information about the scaling parameters and performance enhancements are provided in Supplementary Tables 13 and Supplementary Figure 10.
This improvement to some extent underscored our premise that CO$_2$RR intermediates typically exhibit chemical similarities encompassing both O- and C-type descriptors.
After taking this dual nature into consideration, our hybrid-descriptor method can evaluate the intermediates adsorption energies more accurately.
Importantly, the introduction of hybrid descriptors is computationally affordable compared to DFT methods, as it involves only a set of linear regressions to identify the optimal mixing ratio, as elaborated in the Methods section.

Expanding upon the revised scaling relations, we introduce an activity volcano plot and a rate-determining step (RDS) volcano plot for the CO$_2$RR process, shown in Figure 3a and Supplementary Figure 11.
Given that HER is a key competing reaction with CO$_2$RR [44], we also incorporated a selectivity volcano plot, presented in Supplementary Figure 12, to offer a comprehensive analysis of catalyst candidates.
In adherence to Peterson and Nørskov's original framework [26], our choice of x and y axes descriptors aligned with Gads *CO and Gads *OH.
Further elaboration on how the limiting potentials were evaluated across the volcano plot can be found in the Methods section.
Among the candidatures explored, Ge@g-C3N4 emerged as the most promising catalyst, demonstrating theoretical activity at a limiting potential of -1.2126 eV.
Notably, the observed selectivity trend paralleled the activity trend, suggesting that catalysts exhibiting high theoretical activity are also resistant to HER side reactions.
The RDS volcano plot (Supplementary Figure 11) indicated protonation of *CO as the rate-determining step for nearly optimal candidates.
Moreover, the activity volcano plot predicted an optimal catalyst featuring a theoretical limiting potential of -1.0516 V, at Gads *CO of 0.0589 eV and Gads *OH of -4.0075 eV, positioned at the right-center of the volcano plot.
Results from RDS volcano plot indicate that protonation of *CO likely serves as the RDS for CO$_2$RR.
Consequently, it should be the focal point for future catalyst designs.
The activity and selectivity volcano plots imply that enhancing theoretical activity could be achieved by decreasing CO affinity while increasing OH affinity with the catalysts.
This result suggests a perspective: weak binding of CO to the catalyst surface might facilitate the rotation of C-O backbone from the “upright” configuration in *CO to the “horizontal” orientation in *CHO [45], and therefore facilitating dynamics and protonation of the *CO species.

### 2024-01-04 paper/main_sections/4.3_validation.tex
temp save: format cleanup

Validation of CNN predictions.

Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes [46-48]. The occlusion technique, widely utilized in Computer Vision [49-51], involves systematically masking sections of the input matrix to assess their impact on the final prediction score. In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis [41]. In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy. A visual representation of this process can be found in Supplementary Figure 21. Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states. Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization. This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

To validate our method, we conducted occlusion experiments on Cr@g-C3N4, Co@g-C3N4 and Cu@g-C3N4. Each of these supported metals possesses 3d subshells with different electronic configurations: half-filled, partially filled and fully filled, allowing us to test the universality of our approach. The results of the occlusion experiment on the supported Cr single atom in Cr@g-C3N4 are given in Figure 4a. The d_(x^2-y^2 ) orbital emerges as dominant in CO interaction. Shielding electrons centered around -2 eV significantly alter the adsorption energy, indicating the potency of this orbital. The real-space wavefunction visualization in Figure 4d corroborates these findings, highlighting the interaction between the d_(x^2-y^2 ) orbital of single metal atom and the adsorbate. This suggests the d_(x^2-y^2 ) orbital of supported Co predominantly influence the interaction with CO, and thus shielding electrons from this orbital markedly disturb the adsorption energy. Additionally, occlusion experiments were performed on the Co d_xy orbital of Co@g-C3N4 (Figure 4b) and the Cu d_xy orbital of Cu@g-C3N4 (Figure 4c). In both cases, the d_xy orbital was identified as the dominant contributor. While identifying the dominant orbital is crucial, these findings also suggest that interactions with adsorbates occurring out-of-plane are influenced by in-plane orbitals as well. Importantly, it becomes evident that d orbitals do not universally impact interactions with adsorbates, emphasizing the need for case-specific discussions.

To validate the chemical relevance of predictions made by CNN model, we conducted COHP analysis, a theoretical bond-detecting tool used to identify bonding and anti-bonding contributions to band-structure energy [52]. In this research, interactions between the supported metal atom and atoms within a 5 Å radius were calculated to understand its bonding environment, as detailed in the Methods section. The results of the occlusion experiment on the Co d_xy orbital of Co@g-C3N4, identified as the dominant orbital via occlusion experiments, were compared with the COHP analysis, as shown in Figure 4g and Figure 4i. The occlusion experiment pinpointed strong metal-adsorbate interactions at energy levels of -1 eV and -6.3 eV, which were confirmed by the COHP analysis. We then investigated the non-dominant Co d_xz orbital, and the COHP analysis highlighted interactions at -1 eV and -6.5 eV, aligning with the findings from the occlusion experiments.

>>>>

% Results and Discussion Part Three: Validation of CNN predictions

\section{Validation of CNN predictions}
Historically, interpretability has posed challenges for deep learning methods, often rendering their inner workings akin to black boxes [46-48].
The occlusion technique, widely utilized in Computer Vision [49-51], involves systematically masking sections of the input matrix to assess their impact on the final prediction score.
In our work, we introduce an orbitalwise occlusion method to discern individual orbital contribution to adsorption energy within the CNN model, drawing inspiration from Fung's pioneer integration of this technique into eDOS analysis [41].
In this work, occlusion experiments involve orbitalwise masking specific eDOS, which are then fed into the CNN model to observe resulting disturbance in adsorption energy.
A visual representation of this process can be found in Supplementary Figure 21. Conceptually, this experiment simulates the effect of shielding electronic states, allowing investigation of the potential contribution to adsorption energy from these states.
Additionally, we attempted to extract the planewave coefficients of Kohn-Sham orbital for wavefunction visualization.
This validation step was undertaken to confirm the spatial distribution of the orbitals pinpointed by the occlusion experiments.

To validate our method, we conducted occlusion experiments on Cr@g-C$_3$N$_4$, Co@g-C$_3$N$_4$ and Cu@g-C$_3$N$_4$.
Each of these supported metals possesses 3d subshells with different electronic configurations: half-filled, partially filled and fully filled, allowing us to test the universality of our approach.
The results of the occlusion experiment on the supported Cr single atom in Cr@g-C$_3$N$_4$ are given in Figure 4a.
The d_(x^2-y^2 ) orbital emerges as dominant in CO interaction.
Shielding electrons centered around -2 eV significantly alter the adsorption energy, indicating the potency of this orbital.
The real-space wavefunction visualization in Figure 4d corroborates these findings, highlighting the interaction between the d_(x^2-y^2) orbital of single metal atom and the adsorbate.
This suggests the d_(x^2-y^2) orbital of supported Co predominantly influence the interaction with CO, and thus shielding electrons from this orbital markedly disturb the adsorption energy.
Additionally, occlusion experiments were performed on the Co d_xy orbital of Co@g-C$_3$N$_4$ (Figure 4b) and the Cu d_xy orbital of Cu@g-C$_3$N$_4$ (Figure 4c).
In both cases, the d_xy orbital was identified as the dominant contributor.
While identifying the dominant orbital is crucial, these findings also suggest that interactions with adsorbates occurring out-of-plane are influenced by in-plane orbitals as well.
Importantly, it becomes evident that d orbitals do not universally impact interactions with adsorbates, emphasizing the need for case-specific discussions.

To validate the chemical relevance of predictions made by CNN model, we conducted COHP analysis, a theoretical bond-detecting tool used to identify bonding and anti-bonding contributions to band-structure energy [52].
In this research, interactions between the supported metal atom and atoms within a 5 Å radius were calculated to understand its bonding environment, as detailed in the Methods section.
The results of the occlusion experiment on the Co d_xy orbital of Co@g-C$_3$N$_4$, identified as the dominant orbital via occlusion experiments, were compared with the COHP analysis, as shown in Figure 4g and Figure 4i.
The occlusion experiment pinpointed strong metal-adsorbate interactions at energy levels of -1 eV and -6.3 eV, which were confirmed by the COHP analysis.
We then investigated the non-dominant Co d_xz orbital, and the COHP analysis highlighted interactions at -1 eV and -6.5 eV, aligning with the findings from the occlusion experiments.

### 2024-01-04 paper/main_sections/6_tail.tex
clean up some signs
temp save: format cleanup

Data Availability
Essential source data required to replicate the study's results–including VASP input files, structure files of SACs, eDOS arrays and the pretrained CNN model–are available on GitHub at https://github.com/DanielYang59/cnn4dos or can be provided by the corresponding authors upon request. Due to their substantial size in terabytes, the complete source data is not currently hosted on publicly available repositories but can also be obtained from the corresponding authors upon request.

Code Availability
Associated computer code for this study can be accessed on GitHub at https://github.com/DanielYang59/cnn4dos or obtained from the corresponding authors upon request.

>>>>

Data Availability
Essential source data required to replicate the study's results-including VASP input files,
structure files of SACs, eDOS arrays and the pretrained CNN model-are available on GitHub
at https://github.com/DanielYang59/cnn4dos or can be provided by the corresponding authors upon request.
Due to their substantial size in terabytes, the complete source data is not currently hosted on publicly available repositories but can also be obtained from the corresponding authors upon request.

Code Availability
Associated computer code for this study can be accessed on GitHub at https://github.com/DanielYang59/cnn4dos,
 or obtained from the corresponding authors upon request.
